Multidimensional analysis and detection of informative features in human brain white matter

### 2021-01-27 reports/manuscript/paper.tex
Add more todo notes

Using data from a previous study of patients with amyotrophic lateral sclerosis (ALS)~\cite{sarica2017corticospinal}, we tested the performance of SGL in a
classification setting. The previous study predicted ALS status with a mean
accuracy of 80\% using a random forest algorithm based on \emph{a priori}
selection of features only within the CST bundle-of-interest.
SGL delivers improved predictive performance, with a cross-validated accuracy of {\alsAccuracy}\% and an area under the receiver operating characteristic curve (ROC AUC) of {\alsRocAuc}, without the need for \emph{a priori} feature engineering (Figure \cref{fig:class-results:probs}). In addition to this
classification performance, SGL also identifies the white matter tracts most
important for ALS classification. The relative importance of white matter
features is captured in the $\beta$ coefficients from \cref{supp:eq:sgl} of
the supplemental material \cite{supplement}.
\Cref{fig:class-results:tract-profiles} depicts these coefficients along the
right CST, plotted over the FA values for the control and ALS subject groups.
We find that SGL selects FA metrics in the corticospinal tract and
particularly in the right corticospinal tract as most important to ALS
classification, confirming previous findings \cite{van2011upper,
toosy2003diffusion, sarica2014tractography, sage2007quantitative,
sage2009quantitative, karlsborg2004corticospinal, ellis1999diffusion,
cosottini2005diffusion, ciccarelli2009investigation, abe2010voxel} and
identifying the portions of the brain that were selected \emph{a priori} in
the previous study from which we obtained the data
\cite{sarica2017corticospinal}.

@@

To evaluate the fit of the model, we used a nested cross-validation
procedure. In this procedure, batches of subjects are held out. For each
batch (or fold), the model is fully fit without this data. Then, once the
parameters are fixed, the model is inverted to predict the ages of held out
subjects based on the linear coeffiecients. This scheme automatically finds
the right level of sparsity and fits the coefficients to the ill-posed linear
model, while guarding against overfitting. SGL accurately predicts the age of
the subjects in this procedure, with a median absolute error of {\whMae},
{\hbnMae}, and {\camcanMae} years for the WH, HBN, and Cam-CAN datasets,
respectively and coefficients of determination $R^2 = $ {\whRsq} , {\hbnRsq},
and {\camcanRsq}, respectively (see \cref{fig:age-results}, top panels). The
predictions for Cam-CAN are competitive with a recent state of the art
prediction \cite{mcpherson2020single}, which used streamline density to
estimate the brain's structural connectivity and achieved $R^2 = 0.63$. The
median absolute errors are also lower than the results of a recent study that
predicted age in a large sample that included the Cam-CAN data, and was based
on diffusion MRI features \cite{Richard2018-ux}.

@@

The method is packaged as open-source software called AFQ-Insight that is
openly available \todo{Ariel: add URL}, and provides a clear API to allow for
extensions of the method. The sofware integrates within a broader automated
fiber quantification software ecosystem: AFQ \cite{yeatman2012tract} and
pyAFQ (https://yeatmanlab.github.io), which extract tractometry data from raw
and processed dMRI datasets, as well as AFQ-Browser, which visualizes
tractometry data and facilitates sharing of the results of dMRI studies
\cite{yeatman2018browser}. To facilitate reproducibility and ease use of the
software, the results presented in this paper are also provided in
\url{https://github.com/richford/AFQ-Insight/tree/master/examples/preprint-notebooks}
as a series of Jupyter notebooks \cite{kluyver2016jupyter}.

>>>>

Using data from a previous study of patients with amyotrophic lateral sclerosis (ALS)~\cite{sarica2017corticospinal}, we tested the performance of SGL in a
classification setting. The previous study predicted ALS status with a mean
accuracy of 80\% using a random forest algorithm based on \emph{a priori}
selection of features only within the CST bundle-of-interest.
SGL delivers improved predictive performance, with a cross-validated accuracy of {\alsAccuracy}\% and an area under the receiver operating characteristic curve (ROC AUC) of {\alsRocAuc}, without the need for \emph{a priori} feature engineering (Figure \cref{fig:class-results:probs}). In addition to this
classification performance, SGL also identifies the white matter tracts most
important for ALS classification. The relative importance of white matter
features is captured in the $\beta$ coefficients from \cref{supp:eq:sgl} of
the supplemental material \cite{supplement}.
\Cref{fig:class-results:tract-profiles} depicts these coefficients along the
right CST, plotted over the FA values for the control and ALS subject groups.
We find that SGL selects FA metrics in the corticospinal tract and
particularly in the right corticospinal tract as most important to ALS
classification, confirming previous findings \cite{van2011upper,
toosy2003diffusion, sarica2014tractography, sage2007quantitative,
sage2009quantitative, karlsborg2004corticospinal, ellis1999diffusion,
cosottini2005diffusion, ciccarelli2009investigation, abe2010voxel} and
identifying the portions of the brain that were selected \emph{a priori} in
the previous study from which we obtained the data
\cite{sarica2017corticospinal}.
\todo{Adam: Add sentence comparing to pure lasso results.}

@@

To evaluate the fit of the model, we used a nested cross-validation
procedure. In this procedure, batches of subjects are held out. For each
batch (or fold), the model is fully fit without this data. Then, once the
parameters are fixed, the model is inverted to predict the ages of held out
subjects based on the linear coeffiecients. This scheme automatically finds
the right level of sparsity and fits the coefficients to the ill-posed linear
model, while guarding against overfitting. SGL accurately predicts the age of
the subjects in this procedure, with a median absolute error of {\whMae},
{\hbnMae}, and {\camcanMae} years for the WH, HBN, and Cam-CAN datasets,
respectively and coefficients of determination $R^2 = $ {\whRsq} , {\hbnRsq},
and {\camcanRsq}, respectively (see \cref{fig:age-results}, top panels). The
predictions for Cam-CAN are competitive with a recent state of the art
prediction \cite{mcpherson2020single}, which used streamline density to
estimate the brain's structural connectivity and achieved $R^2 = 0.63$. The
median absolute errors are also lower than the results of a recent study that
predicted age in a large sample that included the Cam-CAN data, and was based
on diffusion MRI features \cite{Richard2018-ux}.
\todo{Adam: Add sentence comparing to pure lasso results.}

@@

The method is packaged as open-source software called AFQ-Insight that is
openly available \todo{Ariel: add URL}, and provides a clear API to allow for
extensions of the method. The sofware integrates within a broader automated
fiber quantification software ecosystem: AFQ \cite{yeatman2012tract} and
pyAFQ (https://yeatmanlab.github.io), which extract tractometry data from raw
and processed dMRI datasets, as well as AFQ-Browser, which visualizes
tractometry data and facilitates sharing of the results of dMRI studies
\cite{yeatman2018browser}. To facilitate reproducibility and ease use of the
software, the results presented in this paper are also provided in
\url{https://github.com/richford/AFQ-Insight/tree/master/examples/preprint-notebooks}
\todo{Adam: Update notebooks}
as a series of Jupyter notebooks \cite{kluyver2016jupyter}.
